\chapter{Acknowledgement}
%This has to be written.
I'd like to thank:
\begin{itemize}
\item Thomas Möller for the continous support during my entire physics studies. He enabled me to go to the US and almost entirely write this thesis abroad. Thomas is a very thoughtful advisor and I found his farsight in our fruitful discussions very inspiring.
%
\item Christoph Bostedt for our close collaboration and the guidance throughout this entire thesis work. Christoph created a great working environment, whether he was in Japan, Germany, or on home-turf in the US. He is an exceptionally gifted experimental physicist, who I learned much from. He has always time for ideas, which also shows his extrodinary leadership skills.
%
\item Kenneth Fergusson, who is very inspiring to talk to and a great colleague to work with. Ken supported the experiments of this thesis significantly. I enjoy our regular conversations and am proud to call you my ``PhD-buddy''.
%
\item Garth Williams, whom I learned a lot from our discussions. This also significantly contributed to this thesis. Beyond this professional level, I am grateful that all our conversations lead to a great friendship.
%
\item Tais Gorkhover for collaborating closely and giving insights in this topic.
%
\item Thomas's group for the support in experiments but also for thesis discussions. Especially Daniela Rupp has had many insights, whether they were for the experiments or the analysis part. Also Maria Müller and Anatoli Ulmer helped with experimental work and discussions.
%
\item The entire LCLS-team, without their close collaboration this thesis experiment would not have been possible. Many people walked the extra mile to make the thesis experiment work and those are Michele Swiggers, Jean-Charles Castagna, Sebastian Carron, Ankush Mitra, Mike Minitti, Michael Holmes, Timur Osipov, Dipawnita Ray, Alex Wallace, Jeff Aldrich, Gabriel Blaj, John Bozek, Ryan Coffee, Ivan Curiel, Peter Noonan, Oliver Hickmann, Jacek Krzywinski, Chris O'Grady, Ago Marinelli, and Alberto Lutmann.
%
\item Our collaborators from Berkley and the USC for their expertise in cruicial experiments for this thesis. These are in particular from USC: Andrey Vilesov, Charles Bernando, Rico Tanyag, Curtis Jones, Justin Kwok, and Luis Gomez. And from Berkely: Oliver Gessner, Michael Ziemkiewicz, Camila Bacellar, and Adam Chatterley.
%
\item Markus Hantschmann for being a great friend throughout my entire physics studies, without whom I would likely not have worked at SLAC.
%
\item Sarah Kaewert for a wonderful time hanging out together, exploring new things, and being my better half.
%
\item Barbara Klimkeit and Werner Schäfer for being there like family.
%
\item My whole family for their love and who made it so easy for me to return home. Particularly my mom Tina, my brother Flo, and my sister Anna had my back.
%
\end{itemize}